---
title: "Week 05 Homework"
author: "Brian Ritz"
date: "Friday, February 13, 2015"
output: html_document
---

** POISSON REGRESSION AND NEGATIVE BINOMIAL REGRESSION **

Import the file:
```{r}
measles.data<-read.csv(file="MeaslesImmunizationCoverageAndOutbreaks.csv")
measles.data<-as.data.frame(measles.data)
```

We will fit the poisson model-- the link function for a poisson model is the log function.

```{r, echo=FALSE}
measles.poisson.model<-glm(Outbreaks~Coverage,family=poisson,data=measles.data)
names(measles.poisson.model)
summary(measles.poisson.model)
```

We will get the linear predictors from the model object:

```{r}
measles.poisson.model$linear.predictors
```

The linear predictors are the log of the fitted values. Below, we compare the outbreaks with the fitted values.
```{r}
rbind(measles.data$Outbreaks,measles.poisson.model$fitted.values)
```

I will show that the link is logrithmic:

```{r}
rbind(log(measles.poisson.model$fitted.values), measles.poisson.model$linear.predictors)
```

What would happen if the coverage changes by 1%?
```{r}
## coverage change in percentage points
(1-exp(measles.poisson.model$coef[2]*.01))*100
```

** NEGATIVE BINOMIAL REGRESSION **

```{r}
library(MASS)
neg.bin.regression.model <- glm.nb(Outbreaks ~ Coverage, data = measles.data, init.theta = 2934552)
summary(neg.bin.regression.model)
```

The negative binomial regression also finds that coverage is significantly negatively correlated with the frequency of cases of Measles. The coefficient on coverage estimated by the negative binomial model is  -0.7828 -- the same as for the poisson model. The standard error of the poisson estimate for the coefficient on coverage is 0.1963, the negative binomial's standard error is the same--0.1963. The deviances are also roughly equal...just 0.002 difference.

I can conclude that the poisson model is appropriate for this data.
