---
title: "Week09 Homework"
author: "Brian Ritz"
date: "Saturday, March 07, 2015"
output: html_document
---


```{r}
HIV.Protease.Data<-read.csv(file="HIV_Protease.csv",header=TRUE,sep=",")
HIV.Protease.Data<-as.data.frame(HIV.Protease.Data)
head(HIV.Protease.Data)
table(HIV.Protease.Data$CODON_2)
```

Show a histogram of the viral loads of the Dataset when CODON_2 = caa:

```{r, echo=FALSE}
HIV.Protease.Data.caa.VL<-subset(HIV.Protease.Data$VL,HIV.Protease.Data$CODON_2=="caa")
hist(HIV.Protease.Data.caa.VL)
```


Show a histogram of the viral loads of the Dataset when CODON_2 = cag:

```{r, echo=FALSE}
HIV.Protease.Data.cag.VL<-subset(HIV.Protease.Data$VL,HIV.Protease.Data$CODON_2=="cag")
hist(HIV.Protease.Data.cag.VL)
```

We will run a model to determine if a change from caa to cag will play a significant role in the failure of the drug.

First we compare the means:
```{r}
c(mean(HIV.Protease.Data.cag.VL),mean(HIV.Protease.Data.caa.VL))
```
The means look similar, but we will need to run statistical tests to confirm that they are not the same.

Fit a linear model:
```{r}
lm.CODON_2.VL <- lm(VL~CODON_2, data=HIV.Protease.Data)
AIC(lm.CODON_2.VL)
anova(lm.CODON_2.VL)
```

From the anova results we can see that the means do not appear different. This is because the p-value on the CODON\_2 response is greater than our threshhold of .05. This is an indicator that perhaps CODON\_2 does not make a difference in the viral load.


Now we will fit the model with a random effect.
```{r}
library(lme4)

lm.CODON_2.VL.Mixed <- lmer(VL ~ 1 + (1|CODON_2), data=HIV.Protease.Data)
AIC(lm.CODON_2.VL.Mixed)
```

The AIC for the random effect is greater than for the fixed effects model. This is surprising in the context of this assignement, but I conclude that the linear model is a better fit to the data.


Now we will fit a logistic regression:

```{r}
glm.CODON_2.VL <- glm(Resp~VL, family="binomial", data=HIV.Protease.Data)
print(AIC(glm.CODON_2.VL))
matplot(1:length(HIV.Protease.Data$Resp),cbind(HIV.Protease.Data$Resp,glm.CODON_2.VL$fitted.values),pch=16)
```

```{r}
head(cbind(HIV.Protease.Data$Resp,glm.CODON_2.VL$fitted.values))
```


We will fit a model with CODON_2 as a predictor:

```{r}
library(MASS)

# glm.CODON_2.LV <- nlme(Resp~VL+CODON_2, data=HIV.Protease.Data)
glm.CODON_2.LV2 <- glmmPQL(Resp~VL, random=~1|CODON_2, family=binomial, data=HIV.Protease.Data)
# glm.CODON_2.LV <- glm(Resp~VL+CODON_2, family = binomial, data=HIV.Protease.Data)
```

Compare the results to other models:
```{r}
AIC(glm.CODON_2.LV2)
plot(predict(glm.CODON_2.LV2,type="response"))
head(cbind(HIV.Protease.Data$Resp,predict(glm.CODON_2.LV2,type="response")))
```

Compare the predictions:

```{r}
par(mfrow=c(1,1))
plot(predict(glm.CODON_2.VL, type="response"))
plot(predict(glm.CODON_2.LV2, type="response"))
```

The plots of the predictions do look different. The model with the random effect of the CODON_2 variable has a concentration of predicitons near zero, which the regular logistic regression does not have. 