---
title: "Week09 Homework"
author: "Brian Ritz"
date: "Saturday, March 07, 2015"
output: html_document
---


```{r}
HIV.Protease.Data<-read.csv(file="HIV_Protease.csv",header=TRUE,sep=",")
HIV.Protease.Data<-as.data.frame(HIV.Protease.Data)
head(HIV.Protease.Data)
table(HIV.Protease.Data$CODON_2)
```

Show a histogram of the viral loads of the Dataset when CODON_2 = caa:

```{r, echo=FALSE}
HIV.Protease.Data.caa.VL<-subset(HIV.Protease.Data$VL,HIV.Protease.Data$CODON_2=="caa")
hist(HIV.Protease.Data.caa.VL)
```


Show a histogram of the viral loads of the Dataset when CODON_2 = cag:

```{r, echo=FALSE}
HIV.Protease.Data.cag.VL<-subset(HIV.Protease.Data$VL,HIV.Protease.Data$CODON_2=="cag")
hist(HIV.Protease.Data.cag.VL)
```

We will run a model to determine if a change from caa to cag will play a significant role in the failure of the drug.

First we compare the means:
```{r}
c(mean(HIV.Protease.Data.cag.VL),mean(HIV.Protease.Data.caa.VL))
```

Fit a linear model:
```{r}
lm.CODON_2.VL <- lm(VL~CODON_2, data=HIV.Protease.Data)
AIC(lm.CODON_2.VL)
anova(lm.CODON_2.VL)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
