---
title: "Workshop1"
author: "Brian Ritz"
date: "Saturday, January 10, 2015"
output: html_document
---
Week 1 Workshop: Comparison of LM() and GLM()
Objective

1. Learn how to fit the classical linear model in the framework of generalized linear model glm().
2. Compare the outputs of lm() and glm() and understand the differences between them.


Load the data to the project:
```{r}
Linear.Model.Data<-read.csv(file="Week1ProjectData.csv",header=TRUE,sep=",")
Linear.Model.Data[1:10,]
Linear.Model.Data.Frame<-as.data.frame(Linear.Model.Data)
```


Now we fit linear model using lm() and glm()

```{r}
Linear.Model.Data.lm<-lm(Output~Input,data=Linear.Model.Data.Frame)
Linear.Model.Data.glm<-glm(Output~Input,family=gaussian(link="identity"),data=Linear.Model.Data.Frame)
```

Names of the lm():
```{r}
names(Linear.Model.Data.lm)
```

Names of the glm()
```{r}
names(Linear.Model.Data.glm)
```

Summary of lm():
```{r}
summary(Linear.Model.Data.lm)
```

Summary of glm():
```{r}
summary(Linear.Model.Data.glm)
```

**COMPARE MODEL OBJECTS**

1. Compare coefficients

Coefficients of the lm():
```{r}
coefficients(Linear.Model.Data.glm)
```

Coefficients of the glm():
```{r}
coefficients(Linear.Model.Data.glm)
```

2. Compare Residuals

Residuals of the lm():
```{r}
residuals(Linear.Model.Data.lm)
```

Residuals of the glm():
```{r}
residuals(Linear.Model.Data.glm)
```

Comparison, check if they are equal:
```{r}
all.equal(residuals(Linear.Model.Data.lm), residuals(Linear.Model.Data.glm))
```

The residuals are EQUAL!

3. Compare fitted values

Fitted values of lm():
```{r}
fitted.values(Linear.Model.Data.lm)
```

Fitted values of glm():
```{r}
fitted.values(Linear.Model.Data.glm)
```

Comparison, check if they are equal:
```{r}
all.equal(fitted.values(Linear.Model.Data.lm), fitted.values(Linear.Model.Data.glm))
```

The fitted values are also EQUAL!


4. Linear Predictions

Linear Predictions for lm():
```{r}
predict(Linear.Model.Data.lm)
```

Linear Predictions for glm():
```{r}
predict(Linear.Model.Data.glm)
```

Comparison of predictions:
```{r}
all.equal(predict(Linear.Model.Data.glm), predict(Linear.Model.Data.lm))
```

The predictions are EQUAL!

5. Deviance

Deviance of lm():
```{r}
deviance(Linear.Model.Data.lm)
```

deviance of glm():
```{r}
deviance(Linear.Model.Data.glm)
```

The deviances are also EQUAL!

NOW, lets calculate the deviance manually:

```{r}


```

