---
title: "Week 04 Homewrok"
author: "Brian Ritz"
date: "Tuesday, February 03, 2015"
output: html_document
---

I chose to study the survival rates of people on the RMS Titanic on April 15, 1912.  April 15 marks the infamous maritime disaster in which over 1,500 people lost their lives in the cold, vast Atlantic Ocean after the Titanic struck an iceburg and disappeared below the waves in the span of just 2 hours.

I will determine the probability of any one passenger surviving as a function of their sex and fare paid for the trip.

I found my data on [Kaggle](https://www.kaggle.com/c/titanic-gettingStarted), a website for data analysis competitions.

```{r}
# import the data from csv
titanic.data <- read.csv("titanic_data.csv", header=TRUE)
# loook at data
dim(titanic.data)
colnames(titanic.data)
class(titanic.data$Sex)
class(titanic.data$Fare)
class(titanic.data$Survived)

```

Make the model, R will create binary variable for us for Sex because it is a factor.

We will fit the models with three different link funcitons, the logit, the probit, and the complementary log-log. All three functions return similar probabilities when the linear predictor is around zero. The complementary log fucntion returns a slightly higher probabilty when the linear predictor is zero.

Towards the extremes, the probit and log-log fucntions return probabilities closer to 1 or 0 faster than the logit function as the linear predictor differs from zero.
```{r}
logit.link.model <- glm(Survived ~ Sex+Fare, data=titanic.data, family=binomial(link="logit"))
probit.link.model <- glm(Survived ~ Sex+Fare, data=titanic.data, family=binomial(link="probit"))
clog.link.model <- glm(Survived ~ Sex+Fare, data=titanic.data, family=binomial(link="cloglog"))
```

Look at the coefficients for each model:
```{r}
data.frame(logit=coef(logit.link.model),probit=coef(probit.link.model),clog=coef(clog.link.model))
```

All three functions return coefficients on the same order of magnitude for each predictor. 

Look at the predicted probabilities:
```{r}
predicted.vals<-data.frame(logit=fitted.values(logit.link.model), probit=fitted.values(probit.link.model), clog=fitted.values(clog.link.model))

predicted.vals[1:10,]
```

Most of the predicted values are very close to eachother. The complementary log-log model appears a little further away from the probit and logit models however.

Lets take a closer look at the logit model:
```{r}
summary(logit.link.model)
exp(coef(logit.link.model)["Sexmale"])
```

Here we see that sex and Fare both play somewhat of a role in predicting whether a passenger will survive the Titanic catastrophe. I conclude that being male made it less likely to survive the disaster. Specifically, being a male reduced your chances of surviving to `r exp(coef(logit.link.model)["Sexmale"])` of what it would be if you were female. Likewise, for every dollar increase in your fare, you were `r exp(coef(logit.link.model)["Fare"])` times as likely to survive.
