---
title: "Week 03 Homework"
author: "Brian Ritz"
date: "Monday, May 04, 2015"
output: html_document
---


The Excel data file is attached.

Use datasets from 1955 to 1968 to build an ARMA or ARIMA models for UN and GDP.
Justify why you chose (ARMA or ARIMA) one over the other.  Note there will be 2 models, one for UN and another for GDP.
Use the chosen UN and GDP models to forecast the UN and the GDP for 1969.
Compare your forecasts with the actual values using error = actual - estimate and plot the errors.
Calculate the Sum of squared(error) for each UN and GDP models.
Regression - build regression models that use:

UN as the independent variable and GDP as the dependent variable - use data from 1955 to 1968 to build the model. Forecast for 1969 and plot the errors and calculate the sum of squared(error) as previously.
GDP as the independent variable and UN as the dependent variable - use data from 1955 to 1968 to build the model. Forecast for 1969 and plot the errors and calculate the sum of squared(error) as previously.
Compare the 2 models - any reason to believe which should be the independent and the dependent variables.

```{r}
library(tseries)
```


Import the data first
```{r}
# import the data

raw.data <- read.csv("Unemployment_GDP_UK.csv", header=T)
unemployment <- ts(data=raw.data$UN, frequency=4, start=c(1955, 2))
gdp <- ts(data=raw.data$GDP, frequency=4, start=c(1955, 2))

```

Use datasets from 1955 to 1968 to build an ARMA or ARIMA models for UN and GDP.

UN Model

Test for stationarity -- Augmented Dickey Fuller Test
```{r}
adf.test(unemployment)
adf.test(gdp)
```

