---
title: "week04_homework"
author: "Brian Ritz"
date: "May 12, 2015"
output: html_document
---

```{r}
require(gdata)
library(lubridate)
library(forecast)
library(dplyr)
# import all the data

date.char <- "July-1"



import.day.data <- function(date.char){
    
  a <- read.xls(paste0("../input/I-57-2013-", date.char, ".xls"), header=T, skip=1, stringsAsFactors=F)
  
  #cnames<-names(a)
  #first.row <- as.character(as.vector(a[1,]))
  
  actual.data <- a[3:26,c(3,5)]
  date.vector <- rep(date.char, nrow(actual.data))
  actual.data$date <- date.vector
  return (actual.data)
}

all.dates<- c("June-16", "June-17", "June-18", "June-19", "June-20", "June-21", "June-22", "June-23", "June-24", "June-25", "June-26", "June-27", "June-28", "June-29", "June-30", "July-1")

all.data<-sapply(all.dates, FUN=import.day.data, simplify=F)

all.data<-do.call(rbind, all.data)

all.data$date <- sapply(all.data$date, FUN=function(x) paste0(x, "-2013"))
all.data$Time <- sapply(all.data$Time, FUN=function(x) paste0(x, ":00"))
all.data$DateTime <- mdy_hms(mapply(FUN=function(x, y) (paste(x, y, sep=" ")), all.data$date, all.data$Time, USE.NAMES=F), tz="America/Chicago")
all.data$I80E <- as.numeric(all.data$I80E)
saveRDS(all.data, file="I80_traffic.Rds")

```

```{r}
# make the time series and fit using auto-arima

I80.traffic<-ts(as.numeric(all.data$I80E), frequency = 24)

june.traffic <- window(I80.traffic, start=c(1,1), end=c(15, 24))
july.traffic <- window(I80.traffic, start=c(16,1), end=c(16, 24))

fit <- auto.arima(june.traffic)


```

DO AICc and BIC select the same model as the best model?

```{r}

summary(fit <- auto.arima(june.traffic, ic="aicc", seasonal = F))
summary(fit <- auto.arima(june.traffic, ic="bic", seasonal=F))



```
Evaluating using AICc or BIC does make a difference. Using AICc, auto.arima produced an arima of (2,0,3), while BIC produced a model of order (2,0,2). 

Instead of changing p and q manually and checking the AICc and BIC of the resulting fit, I used auto.arima to iterate through many p and q parameters and pick the best one based on AICc (in the first case) and BIC (in the second case). 

**PART 2**
Use the day of week seasonal ARIMA(p,d,q)x(P,D,Q)s to forecast July 1, which is Monday.

```{r}

# make a weekly time series
weekly.time.series <- ts(june.traffic, frequency=24*7)

# fit the time series, the seasonality component defaults to the frequency of the time series
summary(weekly.fit<- auto.arima(weekly.time.series, seasonal=T))

```

Now forecast for July 1:
```{r}
plot(july.1.forecast <- forecast(weekly.time.series, h = 24))
summary(july.1.forecast)
```


** PART 3 **
Hour of the day seasonality included:
```{r}
summary(day.seasonality.fit <- auto.arima(june.traffic), seasonal=T)
```
The ARIMA(2,0,1)(2,0,0)[24] tells us that the frequency of the seasonality is 24 hours, or a day. This is daily seasonality. 

July Forecast for the daily seasonality:
```{r}
july.1.forecast.daily <- forecast(day.seasonality.fit, h=24)

(july.1.forecast.daily.means <- july.1.forecast.daily$mean)

```

The forecast for July 1 at 8:00 is `july.1.forecast.daily.means[8]`, the forecast for 9:00 is `july.1.forecast.daily.means[9]`, the forecast for 17:00 is `july.1.forecast.daily.means[17]`, the forecast for 18:00 is `july.1.forecast.daily.means[18]`.


** PART 4 **
For the July 1 8:00, 9:00, 17:00 and 18:00 forecasts, which model is better (part 2 or part 3) ?

```{r}
actuals <- july.traffic
weekly.forecast <- july.1.forecast$mean
daily.forecast <- july.1.forecast.daily.means

weekly.error <- (as.vector(weekly.forecast) - as.vector(actuals))
daily.error <-  (as.vector(daily.forecast) - as.vector(actuals)) 
error.comparison <- weekly.error / daily.error

(error.comparison[8])
(error.comparison[9])
(error.comparison[17])
(error.comparison[18])
```

For all of 8:00, 9:00, 17:00, and 18:00, the weekly forecast performed better than the daily forecast by a wide margin--mostly greater than 10 to 1. 

