---
title: "Course Project Part 2"
author: "Brian Ritz"
date: "Saturday, February 21, 2015"
output: html_document
---


```{r}
Part2.Data<-read.csv(file="../input/MScA_LinearNonLinear_CourseProject_Part2.csv")
Part2.Data<-as.data.frame(cbind(Part2.Data,Part2.Data[,2]/60))
colnames(Part2.Data)<-c("Times","Counts","Temperatures","Intensities")
head(Part2.Data)
```

Visualize the data
```{r}
plot(Part2.Data$Temperatures,Part2.Data$Intensities)
```

**Interpret the plot. What type od relationship you observe?**
I observe a comonotonic relationship -- as temperatures increase, so does the temperature. 

Empirical copula should be normalized-- divide all by n, so all numbersare between 0 and 1.
```{r}
plot(rank(Part2.Data$Temperatures),rank(Part2.Data$Intensities))
```

We see comonotonic behavior, and tail dependence -- pinched corners. We should investigate this. 
The copula can show stratified parts of the model. 
Use clayton because left tail dependence
gumball is for right tail dependence
```{r}
library(MASS)
```

Look at the histogram
```{r}
hist(Part2.Data$Temperatures)
```

Looks normal -- fit dist normal distribution , check kolmogorov - smirnov test -- no reason to reject it is normal  with mean 100.07 and sd 4.812


WE now how two distributions -- one is gamma, one is normal ... fit copula and simulate now.

Output of pobs goes in to fit copula function.

create new object using same parameters for simulating the model.
1. create copula
2. run fit copula -- this estimates the parameters
3. plot the fit copula result

Check out lecture 3. Invert the marginal distributions. 


concentrate on cloud in upper right -- fit differnt model to tail events --see how different they will be
WE first need to separate thet ail from the main sample. Not a lot of obs here, so didn't bother. Probably wont pollute the main sample. 

We separate tail events, plot them, and fit a model to just the tail events.

first part is negative binomial model -- we found earlier that intensity has gamma distribution , so a count has a negative binomial model to original sample.

We tried negative binomial model to tail events, but we wil see what happened. Tail events may come much closer to poisson -- pretty common in practice.  This is because they are significantly rare, and rare events follow a poisson more often. 

Theta comes out of the model -- it is estimated from the long sample 86.3. The regular sample was 4. Significant difference between the two thetas.

What is theta? this is  glm.nb defines the variance of the model as μ+μ2/θ. Theta tells us about overdispersion. Overdispersion is lower when theta is larger. 

Recommended to go and try to fit poisson. 82 deviance vs 83 deg of freedom tells us it is close to poisson.
