---
title: "Week 05 Homework"
author: "Brian Ritz"
date: "Monday, May 18, 2015"
output: html_document
---


Load data from TSA package (the package is written by textbook authors Jonathan Cryer and Kung-Sik Chan).

>library("TSA")

>data(beersales)

The data is the monthly beer sales in millions of barrels, 01/1975 - 12/1990.

Part 1 - use ARIMA(p,d,q) model to forecast beer sales for all months of 1990.

1A - Use the h-period in forecast() to forecast each month of 1990.

1B - Use the monthly data as a continuous time series. Forecast for 1990 Jan, Plug forecast into the time series to forecast for 1990 Feb. And so on and so forth.

1C - which of the two above approaches yield the better results in terms of Mean Squared Error 1990?

Part 2 - use month of the year seasonal ARIMA(p,d,q)(P,Q,D)s model to forecast beer sales for all the months of 1990.

Part 3 - Which model (Part 1 or Part 2) is better to forecast beer sales for each month of 1990 (Jan, Feb, ..., Dec) ? 

Due date: 2015 Feb 19th 11:59pm

```{r}
# load the dataset
library(forecast, quietly=T)
library("TSA", quietly=T)
data(beersales)
```

Create an ARIMA Model

```{r}
train <- window(beersales, start=c(1975,1), end=c(1989,12))
beersales.1990 <- window(beersales, start=c(1990, 1), end=c(1990, 12))
```

adf test checks for stationarity

```{r}
print(adf.test(train))
```

We see that it is stationary because the p-value on the adf test is 0.01. Therefore, we can restrict our model to a stationary model.
```{r}
# we have approximation = F to get rid of warnings that the algorithm could not fit a max likelihood model
fit <- auto.arima(train, max.P=0, max.D=0, max.Q=0, stationary = T, seasonal = F, approximation = F)
```


**1A - Use the h-period in forecast() to forecast each month of 1990.**

```{r}
# the h parameter gives the number of periods forward to forecast
(forecast.1990.arima <- forecast(fit, h = 12))
```

**1B - Use the monthly data as a continuous time series. Forecast for 1990 Jan, Plug forecast into the time series to forecast for 1990 Feb. And so on and so forth.**

```{r}
train.loop <- train
forecasts.loop <- c()
h<-0
while(h < 12){
  fit.loop <- auto.arima(train.loop, max.P=0, max.D=0, max.Q=0, stationary = T, seasonal = F, approximation=F)
  one.mo.forecast <- forecast(fit.loop, h = 1)
  
  # log the forecast we just made
  forecasts.loop <- c(forecasts.loop, one.mo.forecast$mean)
  
  # add the forecast to the training dataset
  train.loop <- ts(c(train.loop, one.mo.forecast$mean), start=start(train.loop), frequency=frequency(train.loop))
  
  # use the length of the forecasts to exit the loop if its time
  h = length(forecasts.loop)
}

(forecasts.loop <- ts(forecasts.loop, start=c(1990, 1), frequency=12))

```

**1C - which of the two above approaches yield the better results in terms of Mean Squared Error 1990?**

```{r}
mse <- function(actual, forecast){
  errors <- as.vector(actual) - as.vector(forecast)
  return (sum(errors**2)/length(errors))
}

forecast.error.yearly <- mse(beersales.1990, forecast.1990.arima$mean)
forecast.error.monthly <- mse(beersales.1990, forecasts.loop)

print(forecast.error.yearly)
print(forecast.error.monthly)
```

The monthly continuous estimation gave a lower mean squared error. 

**Part 2 - use month of the year seasonal ARIMA(p,d,q)(P,Q,D)s model to forecast beer sales for all the months of 1990.**

```{r}
summary(seasonality.fit <- auto.arima(train, seasonal = T, stationary = T, approximation=F))
(seasonality.forecast <- forecast(seasonality.fit, h=12))
```


**Part 3 - Which model (Part 1 or Part 2) is better to forecast beer sales for each month of 1990 (Jan, Feb, ..., Dec) ? **

We will check the mse of part 2, and compare it to part 1:
```{r}
forecast.error.seasonality <- mse(beersales.1990, seasonality.forecast$mean)


print(forecast.error.yearly)
print(forecast.error.monthly)
print(forecast.error.seasonality)

```

Using the seasonality decrease the mean squared error by half! The seasonality model is much better for forecasting beer sales for 1990.


Break it down by forecasts for each month:

```{r}
forecasts.and.actuals <- ts(data.frame(actual = beersales.1990, yearly=forecast.1990.arima$mean, monthly=forecasts.loop, seasonal=seasonality.forecast$mean), start=c(1990, 1), frequency=12)

(all.errors <- apply(forecasts.and.actuals, FUN=function(x) x-forecasts.and.actuals[,"actual"], MARGIN=2))
```

The seasonal forecasts are usually always better than the non-seasonal forecasts, with the exception of May, July, and December.

